---
title: "R Notebook"
output: html_notebook
---

# libraries 

```{r}
library(tidyverse)
library(janitor)
library(hms)
library(lubridate)
library(modelr)
library(caret)
library(broom)
library(leaps)
library(glmulti)
library(tidyverse)
library(pROC)
```

# reading in data

```{r}
flights <- read_csv(here::here("data/flights.csv"))
weather <- read_csv(here::here("data/weather.csv"))
weather_alt <- read_csv(here::here("data/export.csv"))
planes <-  read_csv(here::here("data/planes.csv"))
airlines <- read_csv(here::here("data/airlines.csv"))
airports <- read_csv(here::here("data/airports.csv"))

```


# flights

```{r}
flights %>% 
  arrange(sched_dep_time)

flights_2 <- flights %>% 
  rename(flight_duration = air_time) %>% 
  mutate(across(contains("_time"), ~ sprintf("%04d", .)),
         across(contains("_time"), ~ format(strptime(., format="%H%M"), format = "%H:%M")),
         across(contains("_time"), ~ as.POSIXct(., format = "%H:%M", tz = "UTC")),
         across(contains("_time"), ~ as_hms(.)),
         cancelled_flight = if_else(is.na(dep_time), TRUE, FALSE),
         delayed_flight = if_else(dep_delay > 0 | is.na(dep_delay), TRUE, FALSE)) 


# cancelled flights (ASSUMPTION)

flights_2 %>% 
  filter(cancelled_flight == TRUE)

flights_2 %>% 
  filter(origin == "EWR") %>%
  mutate(weekday = weekdays(time_hour)) %>% 
  mutate(across(weekday, factor, levels=c("Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ggplot(aes(x=weekday, y = dep_delay))+
  geom_boxplot()

flights %>% 
  distinct(dest)


```

```{r}
flights_alt <- flights %>% 
  rename(flight_duration = air_time) %>% 
  mutate(across(contains("_time"), ~ sprintf("%04d", .)),
         across(contains("_time"), ~ format(strptime(., format="%H%M"), format = "%H:%M")),
         across(contains("_time"), ~ as.POSIXct(., format = "%H:%M", tz = "UTC")),
         across(contains("_time"), ~ as_hms(.)),
         cancelled_flight = if_else(is.na(dep_time), TRUE, FALSE),
         delayed_flight = if_else(dep_delay > 0 | is.na(dep_delay), TRUE, FALSE),
         delayed_flight_over30 = if_else(dep_delay > 30 | is.na(dep_delay), TRUE, FALSE)) %>% 
  mutate(date = lubridate::make_date(year = year, month = month, day = day))
```


# weather

```{r}
weather_2 <- weather %>% 
  select(origin, wind_dir, wind_speed, wind_gust, visib, time_hour) %>% 
  fill(contains("wind")) %>% 
  fill(visib) %>%
  drop_na(wind_dir) %>% 
  mutate(wind_dir2 = case_when(
    wind_dir == 0 ~ "No Wind",
    wind_dir <= 22.5 | wind_dir >= 337.5 ~ "North",
    wind_dir > 22.5 & wind_dir < 67.5 ~ "North East",
    wind_dir >= 67.5 & wind_dir <= 112.5 ~ "East",
    wind_dir > 112.5 & wind_dir < 157.5 ~ "South East",
    wind_dir >= 157.5 & wind_dir <= 202.5 ~ "South",
    wind_dir > 202.5 & wind_dir < 247.5 ~ "South West",
    wind_dir >= 247.5 & wind_dir <= 292.5 ~ "West",
    wind_dir > 292.5 & wind_dir < 337.5 ~ "North West",
    TRUE ~ "Other"
  ))

weather_2 %>% 
  count(wind_dir2)

weather %>% 
  count(wind_dir)


# not enough data for temp, dewp, humid, precip and pressure 
# fillforward imputation on wind params
```

```{r}

weather_alt_clean <- weather_alt %>% 
  select(-wpgt, -tsun) %>% 
  drop_na() %>% 
  rename(wind_dir = wdir) %>% 
  mutate(wind_dir2 = case_when(
    wind_dir <= 22.5 | wind_dir >= 337.5 ~ "North",
    wind_dir > 22.5 & wind_dir < 67.5 ~ "North East",
    wind_dir >= 67.5 & wind_dir <= 112.5 ~ "East",
    wind_dir > 112.5 & wind_dir < 157.5 ~ "South East",
    wind_dir >= 157.5 & wind_dir <= 202.5 ~ "South",
    wind_dir > 202.5 & wind_dir < 247.5 ~ "South West",
    wind_dir >= 247.5 & wind_dir <= 292.5 ~ "West",
    wind_dir > 292.5 & wind_dir < 337.5 ~ "North West",
    TRUE ~ "Other"
  )) 


```


# planes

```{r}
planes_2 <- planes %>% 
  select(-speed) %>% 
  mutate(year = na_if(year, 0),
         engines = as.character(engines)) %>% 
  mutate(seats = case_when(
    seats < 11 ~ "10 or less",
    seats < 101 ~ "11 to 100",
    seats < 251 ~ "101 to 250",
    TRUE ~ "Over 250"
  )) %>% 
  rename(year_manufactured = year)
  
planes %>% 
  select(-speed) %>% 
  mutate(year = na_if(year, 0)) %>%
  mutate(seats = case_when(
    seats < 11 ~ "10 or less",
    seats < 101 ~ "11 to 100",
    seats < 251 ~ "101 to 250",
    TRUE ~ "Over 250"
  )) %>% 
  count(seats)
```

# airlines

```{r}
airlines
```

# airports

```{r}
airports %>% 
  ggplot(aes(lon, lat)) +
  geom_point()
```







# flights and airlines join

```{r}
flights_airlines <- flights_2 %>% 
  left_join(airlines, by = "carrier") 

flights_airlines %>% 
  group_by(name) %>% 
  mutate(delayed = if_else(dep_delay > 0, TRUE, FALSE)) %>% 
  summarise(total_flights = n(),
            delayed_flights_pc = sum(delayed, na.rm = TRUE)/n())

flights_airlines %>% 
  filter(name == "Delta Air Lines Inc.",
         origin == "EWR") %>% 
  group_by(month, day) %>% 
  summarise(delayed_flight_pc = sum(delayed_flight, na.rm = TRUE)/n())
  


```



# join everything

```{r}
flights_joined <- flights_2 %>% 
  left_join(airlines, by = "carrier") %>% 
  left_join(weather_2, by = c("origin", "time_hour")) %>% 
  left_join(planes_2, by = c("tailnum")) %>%
  mutate(across(type:engine, ~ case_when(
    is.na(.) ~ "Unknown",
    TRUE ~ .
  ))) %>% 
  mutate(manufacturer = case_when(
    manufacturer == "AIRBUS INDUSTRIE" ~ "AIRBUS",
    manufacturer == "EMBRAER S A" ~ "EMBRAER",
    manufacturer %in% c("AIRBUS", "BOEING", "BOMBARDIER INC", "EMBRAER", "Unknown") ~ manufacturer,
    TRUE ~ "Other"
  )) 





# manufacturers under 10000 put into other cat
# combine airbus and airbus industrie and embraer and embraer S A
```


# first model

## correlations

```{r}
flights_model_data <- flights_joined %>% 
  filter(origin == "EWR",
         cancelled_flight == FALSE) %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -cancelled_flight, -wind_dir, -model, -dest, -delayed_flight) %>% 
  select(dep_delay, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month),
         sched_dep_time = as.numeric(sched_dep_time)) %>% 
 drop_na()


flights_model_data %>% 
  select(dep_delay, 2:6) %>% 
  GGally::ggpairs()

flights_model_data %>% 
  select(dep_delay, 7:11) %>% 
  GGally::ggpairs()

flights_model_data %>% 
  select(dep_delay, 12:15) %>% 
  GGally::ggpairs()
```

```{r}
flights_model_data %>% 
  count(wind_dir2)
```


```{r}
model_1 <- lm(dep_delay ~ wind_gust + wind_dir2 + visib + wind_gust:wind_dir2, flights_model_data)

summary(model_1)
```



# logistic regression

```{r}
flights_model_data_lg <- flights_joined %>% 
  filter(origin == "EWR") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight, -wind_dir, -model, -dest, -dep_delay, -wind_speed, -engines, -engine, -type, - flight_duration) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month),
         sched_dep_time = as.numeric(sched_dep_time)) %>% 
 drop_na()

flights_model_data_lg %>% 
  filter(cancelled_flight == TRUE)
```

```{r}
n_data <- nrow(flights_model_data_lg)

test_index <- sample(1:n_data, size = n_data * 0.2)

test <- flights_model_data_lg %>%  slice(test_index)

train <- flights_model_data_lg %>% slice(-test_index)
```


```{r}
lg_model_1 <- glm(cancelled_flight ~ wind_gust + wind_dir2 + visib + wind_gust:wind_dir2, data = train, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train, family = binomial(link = 'logit'))
summary(lg_model_2)
```


```{r}
predict_log_1 <- test %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test %>% 
  add_predictions(lg_model_2, type = "response")
```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)

```





```{r}
flights_model_data_lg %>% 
  ggplot(aes(delayed_flight, wind_gust))+
  geom_boxplot()
```



# wind speed/dir clusters? 


```{r}
flights_model_data_lg %>% 
  group_by(wind_dir2) %>% 
  summarise(percent_delayed = sum(delayed_flight) / n(),
            avg_wind_speed = mean(wind_gust))
```


```{r}
flights_joined %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x = wind_gust, y = dep_delay)) +
  geom_point()
```

# join alt weather

```{r}
flights_joined_alt <- flights_alt %>% 
  left_join(airlines, by = "carrier") %>% 
  left_join(weather_alt_clean, by = c("date")) %>% 
  left_join(planes_2, by = c("tailnum")) %>%
  mutate(across(type:engine, ~ case_when(
    is.na(.) ~ "Unknown",
    TRUE ~ .
  ))) %>% 
  mutate(manufacturer = case_when(
    manufacturer == "AIRBUS INDUSTRIE" ~ "AIRBUS",
    manufacturer == "EMBRAER S A" ~ "EMBRAER",
    manufacturer %in% c("AIRBUS", "BOEING", "BOMBARDIER INC", "EMBRAER", "Unknown") ~ manufacturer,
    TRUE ~ "Other"
  )) 
```


# lgr different weather data


```{r}
flights_model_data_lg_alt <- flights_joined_alt %>% 
  filter(origin == "EWR") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -delayed_flight_over30) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month),
         sched_dep_time = as.numeric(sched_dep_time)) %>% 
  drop_na()
 

flights_model_data_lg_alt %>% 
  filter(cancelled_flight == TRUE)
```

```{r}
n_data <- nrow(flights_model_data_lg_alt)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_2 <- flights_model_data_lg_alt %>%  slice(test_index)

train_2 <- flights_model_data_lg_alt %>% slice(-test_index)
```


```{r}
lg_model_1 <- glm(cancelled_flight ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_2, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train_2, family = binomial(link = 'logit'))
summary(lg_model_2)
```

```{r}
predict_log_1 <- test_2 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_2 %>% 
  add_predictions(lg_model_2, type = "response")
```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)

```


```{r}
threshold <- 0.04
predict_log_2 <- predict_log_2 %>%
  mutate(pred_thresh_0.6 = pred >= threshold)

head(predict_log_2, 10)

conf_table <- predict_log_2 %>%
  tabyl(cancelled_flight, pred_thresh_0.6)

conf_table
```

#glmulti on cancellations model


```{r}
glmulti_search_all_mains <- glmulti(
  cancelled_flight ~ ., 
  data = train_2,
  level = 1,
  minsize = 4,
  maxsize = 5,
  method = "h",            
  crit = "bic",            
  confsetsize = 10,        
  plotty = F, 
  report = T,              
  fitfunction = "glm",    
  family = binomial(link = "logit")) 
summary(glmulti_search_all_mains)

glmulti_search_previous_mains_one_pair <- glmulti(
  cancelled_flight ~ 1+month+name+sched_dep_time+prcp+wspd, 
  data = train_2,
  level = 2,               
  method = "h",            
  crit = "bic",            
  confsetsize = 10,        
  marginality = TRUE,      
  minsize = 5,            
  maxsize = 6,             
  plotty = F, 
  report = T,              
  fitfunction = "glm",     
  family = binomial(link = "logit")) 

summary(glmulti_search_previous_mains_one_pair)

```

```{r}
lg_model_3 <- glm(cancelled_flight ~ 1+month+name+sched_dep_time+prcp+wspd, data = train_2, family = binomial(link = 'logit'))
summary(lg_model_3)
```

```{r}
predict_log_1 <- test_2 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_2 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_2 %>% 
  add_predictions(lg_model_3, type = "response")
```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```


# lgr delays over 30 mins


```{r}
flights_model_data_lg_alt_2 <- flights_joined_alt %>% 
  filter(origin == "EWR") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -cancelled_flight, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -delayed_flight) %>% 
  select(delayed_flight_over30, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month),
         sched_dep_time = as.numeric(sched_dep_time)) %>% 
  drop_na()
 

flights_model_data_lg_alt %>% 
  filter(cancelled_flight == TRUE)
```

```{r}
n_data <- nrow(flights_model_data_lg_alt_2)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_3 <- flights_model_data_lg_alt_2 %>%  slice(test_index)

train_3 <- flights_model_data_lg_alt_2 %>% slice(-test_index)
```

```{r}
lg_model_1 <- glm(delayed_flight_over30 ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(delayed_flight_over30 ~ ., data = train_3, family = binomial(link = 'logit'))
summary(lg_model_2)

lg_model_3 <- glm(delayed_flight_over30 ~ 1+month+name+sched_dep_time+prcp+wspd, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_3)
```

```{r}
predict_log_1 <- test_3 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_3 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_3 %>% 
  add_predictions(lg_model_3, type = "response")

```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = delayed_flight_over30,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = delayed_flight_over30,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = delayed_flight_over30,
      predictor = pred)



roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```

# better method for sched take off time?

```{r}
flights_joined_alt %>% 
  filter(origin == "EWR") %>% 
  ggplot(aes(x= sched_dep_time, y= dep_delay))+
  geom_point()

flights_joined_alt %>% 
  filter(origin == "EWR") %>% 
  mutate(dep_hr = as_hms(round_date(lubridate::as_datetime(sched_dep_time), "hour"))) %>% 
  group_by(dep_hr) %>% 
  summarise(perc_delayed30 = sum(delayed_flight_over30)/n(),
            perc_delayed = sum(delayed_flight)/n(),
            perc_cancelled = sum(cancelled_flight)/n(),
            number_of_flights = n()) %>% 
  ggplot(aes(x= dep_hr))+
  geom_line(aes(y= perc_delayed), colour = "red")+
  geom_line(aes(y= perc_delayed30), colour = "blue")+
  geom_line(aes(y= perc_cancelled))
  
```

```{r}
flights_joined_alt_3 <- flights_joined_alt %>% 
  mutate(dep_hr = as_hms(round_date(lubridate::as_datetime(sched_dep_time), "hour")),
         time_cat = as.numeric(dep_hr),
         time_cat = case_when(
           time_cat < 28800 ~ "Morning",
           time_cat < 57600 ~ "Day",
           TRUE ~ "Evening"
         )
         )

as.tibble(as.numeric(flights_joined_alt$dep_hr))

flights_joined_alt_3 %>% 
  filter(name=="Frontier Airlines Inc.") %>% 
  drop_na()
```

```{r}
flights_model_data_lg_alt_3 <- flights_joined_alt_3 %>% 
  filter(origin == "EWR") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight_over30, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -dep_hr, -sched_dep_time, -date, - delayed_flight) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month)) %>% 
  drop_na()
 

flights_model_data_lg_alt %>% 
  filter(cancelled_flight == TRUE)
```

```{r}
n_data <- nrow(flights_model_data_lg_alt_3)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_3 <- flights_model_data_lg_alt_3 %>%  slice(test_index)

train_3 <- flights_model_data_lg_alt_3 %>% slice(-test_index)
```

```{r}
lg_model_1 <- glm(cancelled_flight ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train_3, family = binomial(link = 'logit'))
summary(lg_model_2)

lg_model_3 <- glm(cancelled_flight ~ 1+month+name+time_cat+prcp+wspd, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_3)
```

```{r}
predict_log_1 <- test_3 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_3 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_3 %>% 
  add_predictions(lg_model_3, type = "response")

```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = cancelled_flight,
      predictor = pred)



roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```


# previous flight delayed


```{r}
flights_model_data_lg_alt_3 <- flights_joined_alt_3 %>% 
  filter(origin == "EWR") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -dep_hr, -sched_dep_time, -date, -delayed_flight_over30) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month)) %>% 
  drop_na() %>%
  group_by(name) %>% 
  mutate(prev_flight_cancelled = if_else(lag(cancelled_flight) == TRUE, TRUE, FALSE)) %>% 
  drop_na()
```

```{r}
n_data <- nrow(flights_model_data_lg_alt_3)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_3 <- flights_model_data_lg_alt_3 %>%  slice(test_index)

train_3 <- flights_model_data_lg_alt_3 %>% slice(-test_index)
```

```{r}
lg_model_1 <- glm(cancelled_flight ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train_3, family = binomial(link = 'logit'))
summary(lg_model_2)

lg_model_3 <- glm(cancelled_flight ~ 1+month+name+time_cat+prcp+wspd+prev_flight_cancelled, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_3)


```

```{r}
predict_log_1 <- test_3 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_3 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_3 %>% 
  add_predictions(lg_model_3, type = "response")

```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = cancelled_flight,
      predictor = pred)



roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```


# other airports

#lga

```{r}
flights_model_data_lg_alt_3 <- flights_joined_alt_3 %>% 
  filter(origin == "LGA") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -dep_hr, -sched_dep_time, -date, -delayed_flight_over30) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month)) %>% 
  drop_na() %>%
  group_by(name) %>% 
  mutate(prev_flight_cancelled = if_else(lag(cancelled_flight) == TRUE, TRUE, FALSE)) %>% 
  drop_na()
```

```{r}
n_data <- nrow(flights_model_data_lg_alt_3)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_3 <- flights_model_data_lg_alt_3 %>%  slice(test_index)

train_3 <- flights_model_data_lg_alt_3 %>% slice(-test_index)
```

```{r}
lg_model_1 <- glm(cancelled_flight ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train_3, family = binomial(link = 'logit'))
summary(lg_model_2)

lg_model_3 <- glm(cancelled_flight ~ 1+month+name+time_cat+prcp+wspd+prev_flight_cancelled, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_3)


```

```{r}
predict_log_1 <- test_3 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_3 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_3 %>% 
  add_predictions(lg_model_3, type = "response")

```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = cancelled_flight,
      predictor = pred)



roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```

#jfk

```{r}
flights_model_data_lg_alt_3 <- flights_joined_alt_3 %>% 
  filter(origin == "JFK") %>% 
  select(-year, -day, -dep_time, -arr_time, -sched_arr_time, -arr_delay, -carrier, -flight, -tailnum, -origin, -distance, -hour, -minute, -time_hour, -delayed_flight, -wind_dir, -model, -dest, -dep_delay, -engines, -engine, -type, -flight_duration, -dep_hr, -sched_dep_time, -date, -delayed_flight_over30) %>% 
  select(cancelled_flight, everything()) %>% 
  mutate(across(is.character, as.factor),
         month = as.factor(month)) %>% 
  drop_na() %>%
  group_by(name) %>% 
  mutate(prev_flight_cancelled = if_else(lag(cancelled_flight) == TRUE, TRUE, FALSE)) %>% 
  drop_na()
```

```{r}
n_data <- nrow(flights_model_data_lg_alt_3)

test_index <- sample(1:n_data, size = n_data * 0.2)

test_3 <- flights_model_data_lg_alt_3 %>%  slice(test_index)

train_3 <- flights_model_data_lg_alt_3 %>% slice(-test_index)
```

```{r}
lg_model_1 <- glm(cancelled_flight ~ tavg + tmin + tmax + prcp + snow + wspd + pres + wind_dir2, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_1)

lg_model_2 <- glm(cancelled_flight ~ ., data = train_3, family = binomial(link = 'logit'))
summary(lg_model_2)

lg_model_3 <- glm(cancelled_flight ~ 1+month+name+time_cat+prcp+wspd+prev_flight_cancelled, data = train_3, family = binomial(link = 'logit'))
summary(lg_model_3)


```

```{r}
predict_log_1 <- test_3 %>% 
  add_predictions(lg_model_1, type = "response")

predict_log_2 <- test_3 %>% 
  add_predictions(lg_model_2, type = "response")

predict_log_3 <- test_3 %>% 
  add_predictions(lg_model_3, type = "response")

```

```{r}
roc_obj_1 <- predict_log_1 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_2 <- predict_log_2 %>% 
  roc(response = cancelled_flight,
      predictor = pred)

roc_obj_3 <- predict_log_3 %>% 
  roc(response = cancelled_flight,
      predictor = pred)



roc_curve <- ggroc(data = list(roc_obj_1, roc_obj_2, roc_obj_3), legacy.axes = TRUE) +
  coord_fixed() +
  labs(y = "sensitivity (TRR)",
       x = "1 - sensititvity(1 -TNR)")

roc_curve

auc(roc_obj_1)
auc(roc_obj_2)
auc(roc_obj_3)

```